Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-22 paper/main.tex
better motivate method section

For the presented model a list of parameters can be seen in table~\ref{table:list-of-parameters} but
the following calculations are equally true for any model with $N_p$ number of parameters.
Consider the case where we filter out a subset $N_g\leq N_p$ of parameters which will be treated
globally, thus with equal numerical value for every cell.
This means that $N_p - N_g$ parameters will be assigned individually for each of $N_c$ number of
cells.
The total amount of parameters in the system is then given by
\begin{equation}
    N_\text{tot} = N_g + N_c (N_p - N_g).
\end{equation}
It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
the size of the parameter space considerably.
The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
with the dimension ie. the total number of parameters $N_\text{tot}$.
Notice that the presented approach reflects our current work but can easily be adapted to other
scenarios where different strategies of parameters assignment are desired, e.g. when considering
two distinct species of cells which should be displayed by 2 parameters in total regardless of the
number of cells.

@@

\subsubsection{Testing the Comparison Algorithms with Division Events}
When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
data and use the generated masks to provide a benchmark of our algorithm.
To do this, we set out to calculate the cost between two successive iteration steps
$t_n$ with $t_{n+1}$.
We expect to obtain an exponential growth in the cost.
This can be explained by the fact that the growth of the total area is exponential in time.
Furthermore, the temporal area difference can be viewed as a derivative of the total area in time,
which is again an exponential function.
Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.
We can see that by accounting for the parental relationship, we get the overall desired effect.
This method allows us to compare results between states that have a different number of cells.
However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
process this mask as well as the comparator, it is much more computationally demanding.
After all, for each comparison step, we need to run a full 3D visualization pipeline.
In cases where the simpler approach of comparing positions of the agents is possible, this method
should be preferred.
\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}

>>>>

In order to properly estimate the parameters of our model on the single-cell level, we need to be
able to take into account any heterogeneity that may be present.
For the particular model chosen here, a list of parameters can be seen in
table~\ref{table:list-of-parameters} but the following calculations are equally true for any
single-cell model with $N_p$ number of parameters.
Consider the case where we identify a subset $N_g\leq N_p$ of parameters which will be treated
globally, thus with equal numerical value for every cell.
This means that $N_p - N_g$ parameters will be assigned individually for each of $N_c$ number of
cells.
The total amount of parameters $N_\text{tot}$ in the system is then given by
\begin{equation}
    N_\text{tot} = N_g + N_c (N_p - N_g).
\end{equation}
where $N_c$ is the number of unique cells which we encounter over the time interval of interest.
It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
the size of the parameter space considerably.
The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
with the dimension ie. the total number of parameters $N_\text{tot}$.
Notice that the presented approach reflects our current work but can easily be adapted to other
scenarios where different strategies of parameters assignment are desired, e.g. when considering
two distinct species of cells which should be described by 2 parameters in total regardless of the
number of cells.

@@

\subsubsection{Testing the Comparison Algorithms with Division Events}
When utilizing a computational model (see section~\ref{sec:computational-model}), we can generate
data and use the generated masks to provide a benchmark of our algorithm.
To do this, we set out to calculate the cost between two successive iteration steps
$t_n$ and $t_{n+1}$.
We expect to obtain an exponential growth in the cost.
This can be explained by the fact that the growth of the total area is exponential in time.
Furthermore, the temporal area difference can be viewed as a derivative of the total area with
respect to time, which is again an exponential function.
Figure~\ref{fig:mask-difference-metric} (F) shows the total cost for this approach.
We can see that by accounting for the parental relationship, we get the overall desired effect.
This method allows us to compare results between states that have a different number of cells.
However, since this algorithm requires us to calculate a cell-mask for each comparison and then also
process this mask as well as the comparator, it is much more computationally demanding.
After all, for each comparison step, we need to run a full 3D visualization pipeline.
In cases where the simpler approach of comparing positions of the agents is possible, this method
should be preferred.
\textbf{[Wir müssen darüber reden, mir ist hier vieles unklar. Es könnte nützlich sein, vor der Application Sektion, eine Zusammenfassung zu setzen.]}
